FMA: A Dataset For Music Analysis

### 2017-04-22 fma.tex
update

\newcommand{\ntracks}{106,574 }
\newcommand{\nartists}{16,341 }
\newcommand{\nalbums}{14,854 }
\newcommand{\ngenres}{161 }
\newcommand{\tduration}{343 }
\newcommand{\aduration}{278 }
\newcommand{\size}{940 }

@@

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset that can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts and favorites; free-form text such as description, biography, tags, and comments; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 songs from 8 major genres, a medium genre-unbalanced set of 25,000 songs from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{https://github.com/mdeff/fma}.

% Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of metadata for artists analysis or song ratings for recommender systems. Existing datasets such as GTZAN, TagATune, and Million Song suffer from the previous limitations. It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.

@@

\begin{table}[t]
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 729 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{https://github.com/mdeff/fma/}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are clickable links to datasets' homepage.
		\item[2] Audio not directly available, can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32kbps, mono mp3.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison with some other datasets.} % Genre classification datasets.
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & [days] \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\ % MagnaTagATune
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{https://github.com/mdeff/fma/}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Scale is the number of samples times the dimensionality. Size is for a (zipped) archive of all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

@@

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services will not disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, released or available to download audio is usually clips of 10 to 30 seconds, sometimes of low quality.
In comparison, FMA is distributed with legally redistributable full-length, high-quality and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}
% TODO table quality

\textbf{Complete.} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}. Basically all the metadata available through the API has been archived. In addition to the basics like title, album or artist names, this includes per-song genres and user data such as per-track/album/artist favorites, play counts and comments.
% which are necessary to evaluate content-based recommender systems.
There is also free-form text such as per-track/album/artist tags, album description and artist biography.
% social such as artist location or website
% images for tracks, albums and artists
% tags are scarce.

@@

\textbf{Future proof and reproducible.} All files and archives are checksummed and hosted in a long-term digital archive.
%\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
Doing so alleviates the risks of tracks to become unavailable, to change without notice or for online services to shut down.
Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines, so that it can be reproduced or extended.
Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.

@@

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality mp3-encoded songs, hand-picked by established audio curators. Each song is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio. Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks \cite{art:MossFMA}.
Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis \cite{art:MossFMA}.

@@

% Collection
As of April 1st 2017, the date of collection, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks). See \figref{fig:growth} to visualize the growth of the dataset. In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.}. Finally, mp3 audio was downloaded over HTTPS.
% Cleaning
Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71. Finally, the license of 2,616 songs prohibited their redistribution, leaving us with \ntracks tracks.

@@

The collected track, album and artist meta-informations\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks. The audio for each track is stored in a file wich name is the track id. All songs are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{genre_distribution.pdf}
	\\ \vspace{1em}
	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
	\caption{(top) Tracks per genre, from 1 to 38,154.
	(bottom) Tracks per top genre (medium), from 21 to 7,103.
	Experimental music is less represented in the curated set.}
	\label{fig:genre_distribution}
	\label{fig:genre_top_distribution}
\end{figure}

@@

The upper part of \tabref{tab:features} lists the (mostly spectral) features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples. Seven statistics were then computed over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
Those 518 pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{See \texttt{features.py} for the code which computed the features.}

@@

% Cover song recognition --> why not, we have track titles
% Lyrics
% Automatic playlist generation
% style?
% Music classification
	% mood classification,
	% instrument recognition / identification
	% music annotation --> temporal labels, e.g. instrument / note

>>>>

\newcommand{\ntracks}{106,574 }
\newcommand{\nartists}{16,341 }
\newcommand{\nalbums}{14,854 }
\newcommand{\ngenres}{161 }
\newcommand{\tduration}{343 }
\newcommand{\aduration}{278 }
\newcommand{\size}{917 }
\newcommand{\weblink}{https://github.com/mdeff/fma}

@@

% about 150-200 words.
\begin{abstract}
We introduce the Free Music Archive (FMA), an open and easily accessible dataset which can be used to evaluate several tasks in music information retrieval (MIR) such as music classification and annotation or recommendation.
While content-based MIR tasks have typically been solved with feature engineering,
%with a two-stage approach: engineered features are extracted from music audio signals, and are then used as input to a regressor or classifier.
the community is becoming interested in feature learning and end-to-end learning. %, i.e. learning from signals to labels
That endeavor is however restrained by the limited availability of large audio datasets.
% The primary goal of this dataset is to enable the training of large models, which require lots of training data, which could potentially avert the semantic gap currently observed between the extracted low-level features such as MFCCs and the high-level targets such as genre.
By releasing the FMA, we hope to foster research which will improve the state-of-the-art in several tasks and hopefully surpass the performance ceiling observed in e.g.\ genre recognition (MGR).
% as it happened in computer vision with the release of ImageNet.
The released data is composed of \ntracks tracks, \nartists artists, \nalbums albums, \ngenres genres, for a total of \tduration days of audio and \size GiB, all under permissive Creative Commons licenses.
It features a collection of metadata such as the song title, album, artist name and genres; user data such as play counts, favorites, and comments; free-form text such as description, biography, and tags; together with full-length and high-quality audio as well as some pre-computed features. We propose a train/validation/test split and three subsets: a small genre-balanced set of 8,000 tracks from 8 major genres, a medium genre-unbalanced set of 25,000 tracks from 16 genres, and a 94 GiB version of the full set with clips trimmed to 30s.
This paper describes the dataset and how it was created, proposes some tasks, and evaluates some baselines for MGR.
Code, data, and usage examples are available at \url{\weblink}.

% Our major goal is to go beyond the existing limitations of available music datasets, which are either the small size of datasets with raw audio tracks, the availability and legality of the music data, or the lack of metadata for artists analysis or song ratings for recommender systems. Existing datasets such as GTZAN, MagnaTagATune, and Million Song suffer from the previous limitations. It is however essential to establish such benchmark datasets to advance the field of music analysis, like the ImageNet dataset which made possible the large success of deep learning techniques in computer vision.

@@

\begin{table}[t]
	\centering
	\begin{threeparttable}
	\begin{tabular}{l@{ }rrcc}
		\toprule
		dataset\tnote{1} & \#clips & \#artists & year & audio \\
		\midrule
		\href{https://staff.aist.go.jp/m.goto/RWC-MDB/}{RWC} \cite{RWC} & 465 & - & 2001 & yes \\
		\href{http://calab1.ucsd.edu/~datasets/cal500/}{CAL500} \cite{cal500} & 500 & 500 & 2007 & yes \\
		\href{http://mtg.upf.edu/ismir2004/contest/tempoContest/node5.html}{Ballroom} \cite{ballroom} & 698 & - & 2004 & yes \\
		\href{https://marsyasweb.appspot.com/download/data_sets/}{GZTAN} \cite{gtzan} & 1,000 & $\sim300$ & 2002 & yes \\
		\href{http://www.cp.jku.at/datasets/musiclef/}{MusiClef} \cite{musiclef} & 1,355 & 218 & 2012 & yes \\
		\href{https://labrosa.ee.columbia.edu/projects/artistid/}{Artist20} \cite{artist20} & 1,413 & 20 & 2007 & yes \\
		\href{http://ismir2004.ismir.net/genre_contest/}{ISMIR2004} & 1,458 & - & 2004 & yes \\
		\href{http://www-ai.cs.uni-dortmund.de/audio.html}{Homburg} \cite{homburg} & 1,886 & 1,463 & 2005 & yes \\  % GarageBand
		103-Artists \cite{103artists} & 2,445 & 103 & 2005 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{Unique} \cite{unique} & 3,115 & 3,115 & 2010 & yes \\
		\href{http://www.seyerlehner.info/index.php?p=1_3_Download}{1517-Artists} \cite{1517artists} & 3,180 & 1,517 & 2008 & yes \\
		\href{http://www.ppgia.pucpr.br/~silla/lmd/}{LMD} \cite{lmd} & 3,227 & - & 2007 & no \\
		\href{http://anasynth.ircam.fr/home/media/ExtendedBallroom}{EBallroom} \cite{extballroom} & 4,180 & - & 2016 & no\tnote{2} \\
		\href{https://labrosa.ee.columbia.edu/projects/musicsim/uspop2002.html}{USPOP} \cite{uspop} & 8,752 & 400 & 2003 & no \\
		\href{http://calab1.ucsd.edu/~datasets/cal10k/}{CAL10k} \cite{cal10k} & 10,271 & 4,597 & 2010 & no \\ % Swat10k
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{TagATune} \cite{magnatagatune} & 25,863\tnote{3} & 230 & 2009 & yes\tnote{4} \\ % MagnaTagATune
		\href{http://jmir.sourceforge.net/index_Codaich.html}{Codaich} \cite{codaich} & 26,420 & 1,941 & 2006 & no \\ % paper 20,849
		\bf \href{\weblink}{FMA} & \bf \ntracks & \bf \nartists & \bf 2017 & \bf yes \\
		\href{http://www.omras2.org/}{OMRAS2} \cite{omras} & 152,410 & 6,938 & 2009 & no \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} & 1,000,000 & 44,745 & 2011 & no\tnote{2} \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} & 2,084,320 & - & 2017 & no\tnote{2} \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		% 103-Artists is also known as DB-L, Extended Ballroom
		\item[1] Names are clickable links to datasets' homepage.
		\item[2] Audio not directly available, can be downloaded from \href{http://www.ballroomdancers.com}{ballroomdancers.com}, \href{https://www.7digital.com}{7digital.com}, \href{https://www.youtube.com}{youtube.com}.
		\item[3] The 25,863 clips are cut from 5,405 songs.
		\item[4] Low quality 16 kHz, 32 kbit/s, mono mp3.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison with some other datasets.} % Genre classification datasets.
	\label{tab:datasets}
\end{table}
% Lists of datasets:
% * http://www.audiocontentanalysis.org/data-sets/
% * A Survey of Evaluation in Music Genre Recognition

\begin{table*}[t]
	\centering
	\begin{threeparttable}
		\begin{tabular}{l@{\hspace{2.7em}}r
			@{\hspace{2.7em}}>{$}r<{$}
			@{}>{${}={}$}c@{}
			>{$}l<{$}
			@{\hspace{2.7em}}>{$}c<{$}
			@{\hspace{2.7em}}c@{\hspace{2.7em}}c}
		\toprule
		dataset & \#samples & \multicolumn{3}{c}{dimensionality} & \multicolumn{3}{c}{size} \\
		\cmidrule{6-8}
				&           & \multicolumn{3}{c}{}               & \text{volume} & [GiB] & \#days \\
		\midrule
		\href{http://mirg.city.ac.uk/codeapps/the-magnatagatune-dataset}{MagnaTagATune} \cite{magnatagatune} &
			25,863 & 29\cdot16\cdot10^3\cdot1 && 4.6\cdot10^5 & 1.2\cdot10^{10} & 3 & 8.7 \\
		\href{https://research.google.com/audioset/}{AudioSet} \cite{audioset} \tnote{1} &
			2,084,320 & 10\cdot44\cdot10^3\cdot2 && 8.8\cdot10^5 & 1.8\cdot10^{12} & 776 & 241 \\
		\href{https://labrosa.ee.columbia.edu/millionsong/}{MSD} \cite{msd} \tnote{2} &
			1,000,000 & 49\cdot27\cdot10^3\cdot2 && 2.6\cdot10^6 & 2.6\cdot10^{12} & 625 & 567 \\
		\href{\weblink}{FMA} \tnote{2} &
			\ntracks & \aduration\cdot44\cdot10^3\cdot2 && 2.4\cdot10^7 & 2.6\cdot10^{12} & \size & \tduration \\
		\midrule
		\href{http://yann.lecun.com/exdb/mnist/}{MNIST} \cite{mnist} &
			70,000 & 28\cdot28\cdot1 && 7.8\cdot10^2 & 5.5\cdot10^{7\phantom0} & 0.011 & - \\
		\href{https://www.cs.toronto.edu/~kriz/cifar.html}{CIFAR-10} \cite{cifar} &
			60,000 & 32\cdot32\cdot3 && 3.0\cdot10^3 & 1.8\cdot10^{8\phantom0} & 0.16 & - \\
		\href{http://www.image-net.org/challenges/LSVRC/2012/}{ILSVRC 2012-17} \cite{imagenet_challenge} \tnote{3,4} &
			1,341,167 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.0\cdot10^{11} & 100 & - \\
		\href{http://www.image-net.org}{ImageNet} \cite{imagenet} \tnote{3} &
			14,197,122 & 482\cdot415\cdot3 && 6.0\cdot10^5 & 8.5\cdot10^{12} & 1000 & - \\
		\bottomrule
	\end{tabular}
	\begin{tablenotes}
		\item[1] \href{https://support.google.com/youtube/answer/6039860}{YouTube recommended} sample rate for music videos. Size in GiB calculated for mp3 encoded at 320 kbit/s.
		\item[2] Average duration and sample rate. Most excerpts from \href{https://www.7digital.com}{7digital} are 30 or 60s and 22 or 44kHz \cite{msd_features}.
		\item[3] Average resolutions. Most applications resize or crop to $256\cdot256$ for a scale of $2.6\cdot10^{11}$ and $2.8\cdot10^{12}$.
		\item[4] Object detection / classification task. It is the most used dataset for image classification.
	\end{tablenotes}
	\end{threeparttable}
	\caption{Comparison between the largest public audio and image datasets. Dimensionality is (length $\cdot$ sample rate $\cdot$ \#channels) for audio and (xdim $\cdot$ ydim $\cdot$ \#channels) for images. Volume is the number of samples times the dimensionality. Size is for a directory containing all \texttt{.mp3} or \texttt{.jpg} in GiB, an indication of the quantity of information affected by the size and diversity of the data. The last column is the number of days necessary to listen to the whole available audio.}
	\label{tab:size}
\end{table*}
% Magnatagature dead links: http://tagatune.org/Datasets.html http://tagatune.org/Magnatagatune.html
% Announcement: https://musicmachinery.com/2009/04/01/magnatagatune-a-new-research-data-set-for-mir/
% Robin: Alors il télécharge aussi des clips audios de 10 secondes, non-compressé et à 48 kHz pour un total de 4 Terabytes! Donc environ 8x la taille de ton dataset il semble.
% uncompressed: 2084320*10*48e3*2*2 / 1024**4 = 3.6
% 320 kbit/s mp3: 2084320*10*320e3 / 8 / 1024**3 = 776

@@

\textbf{High-quality audio.}
\tabref{tab:datasets} shows that while the smaller datasets are usually distributed with audio, often regardless of copyrights, most of the larger do not.
They either contain features derived from the audio or provide links to download the audio from an online service. The problem with the former is that researchers are stuck with the set of chosen features and are prevented to leverage feature learning algorithms or end-to-end learning systems such as DL, while the later  makes no assurance that the files or services will not disappear or change without notice.
Moreover, researchers are wary of proprietary features such as those computed by commercial services like \href{http://the.echonest.com/}{Echonest} and have extracted features by themselves on the MSD by downloading sample audio, a tedious process \cite{msd_features}.
Finally, distributed or downloadable audio are usually clips of 10 to 30 seconds, sometimes of low quality, e.g.\ 32 kbit/s for MagnaTagATune or an average of 104 kbit/s for MSD \cite{msd_features}.
In comparison, FMA comes with legally redistributable full-length, high-quality (263 kbit/s on average) and non-modified audio.
% \todo{See Table ? for a quantitative analysis of quality.}
% Todo table quality

\textbf{Complete.} In addition to audio, the dataset comes with a wealth of metadata, shown in \tabref{tab:metadata}. All the metadata available through the API has been archived. In addition to the basics like title, album or artist names, this includes per-track genres and user data such as per-track/album/artist favorites, play counts and comments.
% which are necessary to evaluate content-based recommender systems.
There is also free-form text such as per-track/album/artist tags, album description and artist biography. To compare, the MSD only provides artist-level metadata and tags \cite{msd}.
% social such as artist location or website
% images for tracks, albums and artists

@@

\textbf{Future proof and reproducible.} All files and archives are checksummed and hosted in a long-term digital archive.
%\todo{have a DOI and are hosted on \href{https://zenodo.org}{zenodo.org}, a long-term digital archive powered by the \href{https://home.cern}{CERN}}.
Doing so alleviates the risks of songs to become unavailable, to change without notice or for online services to shut down.
Moreover, we share all the code used to (i) collect the data, (ii) analyze it, (iii) generate the subsets and splits, (iv) compute the features and (v) test the baselines. The developed code can serve as a starting point for researchers to compute their own features or evaluate their methods.
Finally, anybody can recreate or extend the collection, thanks to public songs and APIs.

@@

The dataset is a dump of the \href{https://freemusicarchive.org/}{Free Music Archive}, an interactive library of high-quality, legal audio downloads directed by \href{https://wfmu.org/}{WFMU}, the longest-running freeform radio station in the United States.
The archive is funded by the New York State Music Fund, the MacArthur Foundation, the National Endowment for the Arts, and from the project's users.
The website provides a large catalog of artists and high-quality mp3-encoded tracks, hand-picked by established audio curators. Each track is legally free to download as artists decided to release their works under a permissive license, most tracks being released under a \href{https://creativecommons.org/}{Creative Commons license} which allows redistribution of the audio. Cheyenne Hohman, the director of the FMA, \href{http://freemusicarchive.org/member/cheyenne_h/blog/100000_SONGS}{announced} in July 2016 that the website had reached the landmark of 100,000 tracks.
Inspired by Creative Commons and the open-source software movement, the FMA provides a legal and technological platform for curators, artists, and listeners to harness the potential of music sharing. While the archive is free and open to anyone, written and audio content is curated, and permission to upload or edit content is granted on an invitation basis \cite{art:MossFMA}.

@@

% Collection
As of April 1st 2017, the date of collection, the online archive largest track id was 155,320, of which 109,727 were valid (the missing 45,594 ids probably correspond to deleted tracks). See \figref{fig:growth} to visualize the growth of the dataset. In addition to per-track metadata, the used hierarchy of 161 genres and extended album (480 not found) and artist (250 not found) information were collected via the available \href{https://freemusicarchive.org/api}{API}.\footnote{See \texttt{webapi.ipynb} to query the API with our helpers.}. Finally, mp3 audio was downloaded over HTTPS.
% Cleaning
Out of all collected track ids, 180 mp3s could not be downloaded, 286 could not be trimmed by ffmpeg, and features could not be extracted from 71. Finally, the license of 2,616 tracks prohibited their redistribution, leaving us with \ntracks tracks.

@@

The collected track, album and artist meta-informations\footnote{\texttt{raw\_tracks.csv}, \texttt{raw\_albums.csv}, \texttt{raw\_artists.csv}} were cleaned, uniformly formatted, merged and stored in \texttt{tracks.csv}\footnote{\label{creation}See \texttt{creation.ipynb} for the code which created the dataset.} which \tabref{tab:tracks} shows an excerpt. That file is a relational table where each row represents a track and columns are listed in \tabref{tab:metadata}, although some values may be missing for some tracks. The audio for each track is stored in a file wich name is the track id. All tracks are mp3-encoded, most of them with sampling rate of 44,100 Hz, bit rate 320 kb/s, and in stereo. Figures~\ref{fig:album_release_year},~\ref{fig:duration_distribution} and \ref{fig:listens_distribution} show the distribution of albums per year, track durations, and play counts per album.
See the \texttt{analysis.ipynb} notebook for a much more detailed analysis of the content.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\linewidth]{genre_distribution.pdf}
	\\ \vspace{1em}
	\includegraphics[width=\linewidth]{genre_top_distribution.pdf}
	\caption{(top) Tracks per genre on the full set, from 1 to 38,154.
	(bottom) Tracks per all top genre on the medium subset, from 21 to 7,103.
	Note how experimental music is much less represented in the curated medium subset.}
	\label{fig:genre_distribution}
	\label{fig:genre_top_distribution}
\end{figure}

@@

\tabref{tab:features} (upper part) lists the features we extracted from the audio with the \href{https://github.com/librosa/librosa}{librosa} Python library, version 0.5.0 \cite{librosa}.
Each feature set (except zero-crossing rate) is computed on windows of 2048 samples spaced by hops of 512 samples. Seven statistics were then computed over all windows: the mean, standard deviation, skew, kurtosis, median, minimum and maximum.
Those 518 pre-computed features are distributed in \texttt{features.csv} for all tracks.\footnote{See \texttt{features.py} for the code which computed the features.}

@@

% Cover song detection / recognition --> why not, we have track titles
% Lyrics
% Automatic playlist generation
% style?
% Music classification
	% mood classification --> may be in tags
	% instrument recognition / identification --> may be in tags
	% music annotation --> temporal labels, e.g. instrument / note
